---
title: "assignment8"
author: "William Hope"
date: "2023-03-27"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(stargazer)
library(knitr)
library(sandwich)
library(lmtest)
library(AER)
library(momentfit)
library(tinytex)
```

```{r functions, echo=FALSE}
getDat4 <- function(n, a1=20, a2=1, a3=-2, d1=100, d2=-1, d3=2, sige=16,
                    sigeta=16, sigX=5, sigZ=5)
{
    e <- rnorm(n, 0, sqrt(sige))
    eta <- rnorm(n, 0, sqrt(sigeta))
    X1 <- rnorm(n, 10, sqrt(sigX))
    X2 <- rnorm(n, 10, sqrt(sigX))
    Z <- rnorm(n, 10, sqrt(sigZ))
    d4 <- runif(1,-.01,.01)
    Q <- (a2*(d1+eta+d3*X1+d4*X2)-d2*(a1+a3*Z+e))/(a2-d2)
    P <- ((d1+eta+d3*X1+d4*X2)-(a1+a3*Z+e))/(a2-d2)
    dat <- data.frame(P,Q,Z,X1,X2,e,eta)
    list(dat=dat, a=c(a1,a2,a3), d=c(d1,d2,d3,d4), sig=c(e=sige, eta=sigeta))
}
plotEqu4 <- function(obj, nCurves=NULL, nPoints=NULL, f=.9, t=1.1)
{
    dat <- obj$dat
    if (!is.null(nPoints))
        dat <- dat[1:nPoints,]
    plot(Q~P, dat, pch=21, col="lightblue", bg="lightblue",
         main="Demands and Supplies with their equilibrium points",
         bty='n')
    if (is.null(nCurves))
    {
        n <- nrow(dat)
    } else {
        n <- nCurves
        if (nCurves>nrow(dat))
            stop("The number of curves cannot exceed the number of points")
    }
    for (i in 1:n)
    {
        curve(obj$d[1]+obj$d[2]*x+obj$d[3]*dat$X1[i]+obj$d[4]*dat$X2[i]+dat$eta[i],
              dat$P[i]*f, dat$P[i]*t,
              col="green", add=TRUE)
        curve(obj$a[1]+obj$a[2]*x+obj$a[3]*dat$Z[i]+dat$e[i],
              dat$P[i]*f, dat$P[i]*t,
              col="orange", add=TRUE)
    }
    points(dat$P[1:n], dat$Q[1:n], pch=23, col="darkred",bg="darkred")
    grid()
}

printEq4 <- function(obj)
{
    cat("\\begin{eqnarray*}\n")
    cat("Q^d &=& ", obj$d[1], ifelse(obj$d[2]<0, "-", "+"), abs(obj$d[2]), "P",
        ifelse(obj$d[3]<0, "-", "+"), abs(obj$d[3]), "X_1",
        ifelse(obj$d[4]<0, "-", "+"), abs(obj$d[4]), "X_2 + \\eta\\\\\n",
        "Q^s &=& ", obj$a[1], ifelse(obj$a[2]<0, "-", "+"), abs(obj$a[2]), "P",
        ifelse(obj$a[3]<0, "-", "+"), abs(obj$a[3]), "Z + e\n",
        "\\end{eqnarray*}\n", sep="")
}  
```

# Part 1

For this part, we need to generate new demand and supply data, using
different functions (hidden above). It looks similar to the previous
assignment, but the demand now depends on two exogenous regressors.

```{r, fig.align='center', out.width='60%'}
set.seed(20885971)
obj <- getDat4(200)
plotEqu4(obj, 5, 20)
```

```{r, results='asis'}
printEq4(obj)
```

## Estimating the Supply function

The model we want to analyze is the following supply and demand system:

\begin{eqnarray*}
Q^d & = & \delta_1 + \delta_2 P + \delta_3 X_1 +\delta_4 X_2+ \eta\\
Q^s & = & \alpha_1 + \alpha_2 P + \alpha_3 Z + e
\end{eqnarray*}

where $X_1$, $X_2$ and $Z$ are exogenous variables (uncorrelated with
$\eta$ and e).

First, we generate a dataset using my student ID as seed by running
the following code.

```{r}
set.seed(20885971) ## Entering student ID
sigeta <- runif(1, 8, 25)
sige <- runif(1, 8, 25)
sigX <- runif(1, 3, 10)
sigZ <- runif(1, 3, 10)
d3 <- round(runif(1,-3,3),2) # The exogenous demand shifter
a3 <- round(runif(1, -3,3),2) # The exogenous supply shifter
d2 <- round(runif(1, -5,-.5),2) # the demand slope
a2 <- round(runif(1, .5, 5),2) # the supply slope
obj <- getDat4(n=100, a2=a2, d2=d2, d3=d3, a3=a3, sige=sige, sigeta=sigeta,
               sigX=sigX, sigZ=sigX)
save(obj, file="A8Data.rda")
load("A8Data.rda")
```

We only want to estimate the supply function. I want you to try the
following sets of excluded exogenous variables: (i) $\{X_1\}$, (ii)
$\{X_2\}$ and (iii) $\{X_1,X_2\}$. Don't forget that the set of
instruments must include the included exogenous variable (Z in this
case). For each set of instruments, answer the following questions.

# Question 1

Estimated the reduced form and testing if the instruments are strong. 

```{r}

red <- lm(P~X1, obj$dat)
knitr::kable(coeftest(red, vcov.=vcovHC)[,])

blue <- lm(P~X2, obj$dat)
knitr::kable(coeftest(blue, vcov.=vcovHC)[,])

yellow <- lm(P~X1+X2+Z, obj$dat)
knitr::kable(coeftest(yellow, vcov.=vcovHC)[,])

```

When testing X1 by itself, it is significant. X2 is not significant.
Morever, when testing all the instrumental variables, X1 and Z are significant.


# Question 2

Using `ivreg` to estimate the model by TSLS. Interpretation below.

```{r}
# Using IRVEG with X1
fit <- ivreg(Q~P | X1, data=obj$dat)
fit_OLS <- lm(Q~P+X1, data=obj$dat) # Using OLS to compare results 

se <- sqrt(diag(vcovHAC(fit)))
se2 <- sqrt(diag(vcovHAC(fit_OLS)))

stargazer(fit, fit_OLS, se=list(se, se2), header=FALSE, float=FALSE, type = 'text')

# Using IRVEG with X2
fit <- ivreg(Q~P | X2, data=obj$dat)
fit_OLS <- lm(Q~P+X2, data=obj$dat) # Using OLS to compare results 

se <- sqrt(diag(vcovHAC(fit)))
se2 <- sqrt(diag(vcovHAC(fit_OLS)))

stargazer(fit, fit_OLS, se=list(se, se2), header=FALSE, float=FALSE, type = 'text')

# Using IRVEG with X1, X2, and Z
fit <- ivreg(Q~P | X1+X2+Z, data=obj$dat)
fit_OLS <- lm(Q~P+X1+X2+Z, data=obj$dat) # Using OLS to compare results 

se3 <- sqrt(diag(vcovHAC(fit)))
se4 <- sqrt(diag(vcovHAC(fit_OLS)))
stargazer(fit, fit_OLS, se=list(se, se2), header=FALSE, float=FALSE, type = 'text')


```

When reviewing the results, it is clear IV does a better job with X1, while OLS
does a better job with the other models. 


# Question 3

Testing if the instruments are valid. Interpretation below.


```{r}

fit <- ivreg(Q~P | X1, data=obj$dat)
AER:::ivdiag(fit)

fit <- ivreg(Q~P | X2, data=obj$dat)
AER:::ivdiag(fit)

fit <- ivreg(Q~P | X1+X2+Z, data=obj$dat)
AER:::ivdiag(fit)


```
We have strong instruments when we use X1, X2, and Z. 

# Question 4

Testing if $P$ is exogenous. Interpretation below.

```{r}

red <- lm(P~X1, obj$dat) # testing exogeneity of price with X1, X2, and Z
uhat <- residuals(red)
cf <- lm(Q~P + uhat, obj$dat)
knitr::kable(coeftest(cf, vcov=vcovHC)[,])

red <- lm(P~X2, obj$dat) # testing exogeneity of price with X1, X2, and Z
uhat <- residuals(red)
cf <- lm(Q~P + uhat, obj$dat)
knitr::kable(coeftest(cf, vcov=vcovHC)[,])

red <- lm(P~X1+X2+Z, obj$dat) # testing exogeneity of price with X1, X2, and Z
uhat <- residuals(red)
cf <- lm(Q~P + uhat, obj$dat)
knitr::kable(coeftest(cf, vcov=vcovHC)[,])

```

It appears P is exogenous when we use X2, but not exogenous in the other cases.

Ideally, it appears that the 1st method is not better among the 3. The reason 
why is because the 2nd method, using AER performs multiple tests and allows you
to decide with fewer lines of code. It is also much easier to read and 
understand whether the IV are significant. The 1st does provide the 
same/similar results, but the 2nd and 3rd methods are much more concise.

# Part 2

For this part, we want to work with the fish data used by Kathryn
Graddy in the paper *Fulton Fish Market* published in the Journal of
Economic Perspectives in 2006. The dataset contains daily data
from December 1991 to March 1992. We use the `read.table` function to extract 
the data. 

Note: I use my specific directory, which must be changed for external 
use to your directory.

```{r}
Fulton <- read.table("C:/Users/savag/OneDrive/Documents/Econ 323/Fulton.txt") 
# change when necessary to your directory
```

## Question 1:

We reproduce the Table 2 from the paper. For the IV regressions (there
are 2 in the table), we test the exogeneity of price and the relevance of
the instrument (Stormy). Interpretation below.

```{r}
fit <- lm(LogQuantity~LogPrice, Fulton)
fit2 <- lm(LogQuantity~LogPrice+Monday+Tuesday+Wednesday+Thursday+Cold+Rainy, Fulton)
fit3 <- ivreg(LogQuantity~LogPrice, data=Fulton)
fit4 <- ivreg(LogQuantity~LogPrice+Monday+Tuesday+Wednesday+Thursday+Cold+Rainy | Cold+Rainy+Monday+Tuesday+Wednesday+Thursday+Stormy, data=Fulton)

se <- sqrt(diag(vcovHAC(fit)))
se2 <- sqrt(diag(vcovHAC(fit2)))
se3 <- sqrt(diag(vcovHAC(fit3)))
se4 <- sqrt(diag(vcovHAC(fit4)))
stargazer(fit, fit2, fit3, fit4, se=list(se, se2, se3, se4), header=FALSE, float=FALSE, type = 'text')

red <- lm(LogPrice~Stormy, Fulton) # testing exogeneity of price
uhat <- residuals(red)
cf <- lm(LogQuantity~LogPrice + uhat, Fulton)
knitr::kable(coeftest(cf, vcov=vcovHC)[,])

# relevance of the instrument (Stormy)
yellow <- lm(LogPrice~Stormy++Monday+Tuesday+Wednesday+Thursday, Fulton) 

knitr::kable(coeftest(yellow, vcov.=vcovHC)[,])

AER:::ivdiag(fit4)

```

When looking at the new table, reproduced from the paper, we can see that the 
coefficients are similar and the standard errors are very similar. While
some are different, it is clear that our tables are the same when calculating
each variable's significance from the paper and comparing the results.


When testing the exogeneity of log price, we see that the coefficient of uhat 
is highly insignificant, so we strongly accept the hypothesis that ed76 is 
exogenous.


## Question 2

We cannot test the validity of the instrument when we only use
$Stormy$. We estimate the demand using the instruments `Cold` and `Rainy`
and perform all tests (validity, relevance and exogeneity). 

```{r}

red <- lm(LogPrice~Cold+Rainy, Fulton)
edHat <- fitted(red)
sec <- lm(LogQuantity~edHat, Fulton)
b <- coef(sec)
uhat <- with(Fulton, LogQuantity-b[1]-b[2]*LogPrice)
J <- lm(uhat~Cold+Rainy, Fulton)
nobs(J)*summary(J)$r.square

qchisq(.95,1)

uhat <- residuals(red)
cf <- lm(LogQuantity~LogPrice + uhat, Fulton)
knitr::kable(coeftest(cf, vcov=vcovHC)[,])

```

Interpretation of the results. Which instrument(s) seems to be the best choice?

We do not reject the hypothesis that the instruments are exogenous.

We can see that the coefficient of uhat is highly insignificant, we strongly 
reject the hypothesis that ed76 is NOT exogenous. 

## Question 3:

We want to see in this question if the demand function can be
estimated. Assume the following demand and supply:

\begin{eqnarray*}
\log(Q^s) &=& \alpha_1 + \alpha_2\log(P)+\alpha_3 Mon + \alpha_4 Tue + 
\alpha_5 Wed + \alpha_6 Thu + \alpha_7 Stormy + e\\
\log(Q^d) &=& \delta_1 + \delta_2\log(P) + \delta_3 Cold + \delta_4 Rainy + u
\end{eqnarray*}

Can we estimate the supply function? We use TSLS and by performing the 
appropriate tests.

```{r}

supplyIV <- ivreg(LogQuantity~LogPrice+Monday+Tuesday+Wednesday+Thursday | Monday+Tuesday+Wednesday+Thursday+Stormy, data=Fulton)
knitr::kable(coeftest(supplyIV, vcov=vcovHC)[,])

AER:::ivdiag(supplyIV)

```

We see that the test results in strong instruments. It appears we can estimate
the supply function.

